### 2019-07-02 scipy-1.0/consortium.tex
Updated author list and contribution statement, added consortium

{\bf
Aditya Vijaykumar$^{31,32}$, Alessandro Pietro Bardelli$^{10}$, Alex Rothberg$^{10}$, Andreas Hilboll$^{33}$, Andreas Kloeckner$^{34}$, Antony Lee$^{35}$, Ariel Rokem$^{36}$, C. Nathan Woods$^{9}$, Chad Fulton$^{37}$, Charles Masson$^{38}$, Christian Brueffer$^{39}$, Christian H\"aggstr\"om$^{40}$, Clark Fitzgerald$^{41}$, David A. Nicholson$^{42}$, David R Hagen$^{43}$, Dmitrii V. Pasechnik$^{44}$, Emanuele Olivetti$^{45}$, Eric Martin$^{10}$, Eric Wieser$^{46}$, Fabrice Silva$^{47}$, Felix Lenders$^{48,49,50}$, Florian Wilhelm$^{51}$, G. Young$^{10}$, Gavin A. Price$^{52}$, Gert-Ludwig Ingold$^{53}$, Gregory E. Allen$^{54}$, Gregory R. Lee$^{55,56}$, Herv\'e Audren$^{57}$, Irvin Probst$^{58}$, J\"org P. Dietrich$^{59,60}$, Jacob Silterra$^{61}$, James T Webber$^{62}$, Janko Slavi\v{c}$^{63}$, Joel Nothman$^{64}$, Johannes Buchner$^{65,66}$, Johannes Kulick$^{67}$, Johannes L. Sch\"{o}nberger$^{10}$, Jos\'e Vin\'icius de Miranda Cardoso$^{68}$, Joscha Reimer$^{69}$, Joseph Harrington$^{70}$, Juan Luis Cano Rodr'iguez$^{71}$, Juan Nunez-Iglesias$^{72}$, Justin Kuczynski$^{73}$, Kevin Tritz$^{74}$, Martin Thoma$^{75}$, Matthias K\"ummerer$^{76}$, Maximilian Bolingbroke$^{77}$, Michael Tartre$^{78}$, Mikhail Pak$^{79}$, Nathaniel J. Smith$^{10}$, Nikolai Nowaczyk$^{10}$, Nikolay Shebanov$^{80}$, Oleksandr Pavlyk$^{81}$, Per A. Brodtkorb$^{82}$, Perry Lee$^{10}$, Robert T. McGibbon$^{83}$, Roman Feldbauer$^{84}$, Sam Lewis$^{85}$, Sam Tygier$^{86}$, Scott Sievert$^{87}$, Sebastiano Vigna$^{88}$, Sourav Singh$^{89}$, Stefan Peterson$^{10}$, Surhud More$^{90,91}$, Tadeusz Pudlik$^{92}$, Takuya Oshima$^{93}$, Thomas J. Pingel$^{94}$, Thomas P. Robitaille$^{95}$, Thomas Spura$^{10}$, Thouis R. Jones$^{96}$, Tim Cera$^{10}$, Tim Leslie$^{10}$, Tiziano Zito$^{97}$, Tom Krauss$^{98}$, Utkarsh Upadhyay$^{99}$, Yaroslav O. Halchenko$^{100}$, Yoshiki V\'azquez-Baeza$^{101}$
}
\newline
\hfill \break
$^{31}$International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bengaluru 560089, India, $^{32}$Department of Physics, Birla Institute of Technology and Science, Pilani 330331, India, $^{33}$Institute of Environmental Physics, University of Bremen, Bremen, Germany, $^{34}$Department of Computer Science, University of Illinois at Urbana-Champaign, 201 N. Goodwin Ave, Urbana, IL 616801, USA, $^{35}$Laboratoire Photonique, Num\'erique et Nanosciences UMR 5298, Universit\'e de Bordeaux, Institut d'Optique Graduate School, CNRS, 33400 Talence, France, $^{36}$The University of Washington eScience Institute, The University of Washington, Seattle, WA 98105, USA, $^{37}$Federal Reserve Board of Governors,  20th and Constitution Ave. NW, Washington, DC 20551, U.S.A., $^{38}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, $^{39}$Department of Clinical Sciences Lund, Lund University, Lund, Sweden, $^{40}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, $^{41}$Statistics Department, University of California - Davis, Davis, CA 95616, USA, $^{42}$Emory University, Atlanta, GA 30322, USA, $^{43}$Applied BioMath, 561 Virginia Rd Ste 220, Concord, MA 01742, USA, $^{44}$Department of Computer Science, University of Oxford, Wolfson Building, Parks Road, OX1 3QD Oxford, UK, $^{45}$NeuroInformatics Laboratory, Bruno Kessler Foundation, Trento 38123, Italy, $^{46}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, $^{47}$Aix Marseille Universit\'e., CNRS, Centrale Marseille, LMA, Marseille, France, $^{48}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, $^{49}$ABB Corporate Research, ABB AG, Wallstadter Str. 59 68526 Ladenburg, Germany, $^{50}$Institut f\"ur Mathematische Optimierung, Technische Universit\"at Carolo-Wilhelmina zu Braunschweig, Universit\"atsplatz 2 38106 Braunschweig, Germany, $^{51}$Independent Researcher, Cologne, Germany, $^{52}$Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA, $^{53}$Institut f{\"u}r Physik, Universit{\"a}t Augsburg, Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany, $^{54}$Applied Research Laboratories, The University of Texas at Austin, P.O. Box 8029 Austin, TX 78713, USA, $^{55}$Department of Radiology, School of Medicine, University of Cincinnati, Cincinnati, OH 45221, USA, $^{56}$Department of Radiology, Cincinnati Children's Hospital Medical Center, Cincinnati, OH 45229, USA, $^{57}$Ascent Robotics Inc., 1-6-10 Hiroo Shibuya Tokyo, Japan, $^{58}$ENSTA Bretagne, 2 rue François Verny, Brest, France, $^{59}$Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 1, 81679 M\"unchen, Germany, $^{60}$Excellence Cluster Universe, Boltzmannstr. 2, 85748 Garching b. M\"unchen, Germany, $^{61}$Independent Researcher, USA, $^{62}$Data Sciences, Chan Zuckerberg Biohub, San Francisco, CA 94158, USA, $^{63}$Faculty of Mechanical Engineering, University of Ljubljana, A\v{s}ker\v{c}eva 6 1000 Ljubljana, Slovenia-EU, $^{64}$Sydney Informatics Hub, The University of Sydney, NSW 2006, Australia, $^{65}$Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'{o}lica de Chile, Casilla 306, Santiago 22, Chile, $^{66}$Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85741 Garching, Germany, $^{67}$University of Stuttgart, Machine Learning and Robotics Lab, Universitätsstr. 38, 70569 Stuttgart, Germany, $^{68}$Department of Electrical Engineering, Universidade Federal de Campina Grande, 58429900 PB, Brazil, $^{69}$Department of Computer Science, Kiel University, 24098 Kiel, Germany, $^{70}$Planetary Sciences Group and Florida Space Institute and Department of Physics, University of Central Florida, 4111 Libra Drive Orlando Florida 32816-2385, USA, $^{71}$Independent Researcher, Spain, $^{72}$Monash Micro Imaging, Monash University, Clayton VIC 3800, Australia, $^{73}$Department of Molecular, Cellular, and Developmental Biology, University of Colorado, Boulder, Boulder, CO 80303, USA, $^{74}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N. Charles Street, Baltimore, MD 21218, USA, $^{75}$Independent Researcher, Munich, Germany, $^{76}$University of T\"ubingen, T\"ubingen, Germany, $^{77}$Independent Researcher, United Kingdom, $^{78}$Two Sigma Investments, 100 Avenue of the Americas 16th Floor, New York, NY 10013, USA, $^{79}$Department of Mechanical Engineering, Technical University of Munich, Boltzmannstr. 15, 85748 Garching, Germany, $^{80}$Independent Researcher, Berlin, Germany, $^{81}$Intel Corp., Austin, TX 78746, USA, $^{82}$Independent Researcher, Horten, Norway, $^{83}$D. E. Shaw Research, New York, NY 10036, USA, $^{84}$Division of Computational Systems Biology, University of Vienna, Althanstra{\ss}e 14, 1090 Vienna, Austria, $^{85}$Independent Researcher, Melbourne, Australia, $^{86}$School of Physics and Astronomy, University of Manchester, Oxford Road, M13 9PL, UK, $^{87}$Electrical and Computer Engineering, University of Wisconsin--Madison, 1415 Engineering Drive Madison, WI, USA, $^{88}$Dipartimento di Informatica, Universit\`a degli Studi di Milano, via Celoria 18, 20133 Milano MI, Italy, $^{89}$Department of Computer Engineering, Vishwakarma Institute of Information Technology, Pune, India, $^{90}$Inter-University Centre for Astronomy and Astrophysics, Ganeshkhind, Pune, 411007, India, $^{91}$Kavli Institute for the Physics and Mathematics of the Universe, Kashiwanoha 5-1-5, Kashiwa-shi, 2778583, Japan, $^{92}$Waymo LLC, Mountain View, CA 94043, USA, $^{93}$Faculty of Engineering, Niigata University, 8050 Ikarashi-Ninocho, Nishi-ku, Niigata, 9502181, Japan, $^{94}$Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA, $^{95}$Aperio Software, Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom, $^{96}$Broad Institute, Cambridge, MA 02142, USA, $^{97}$Department of Psychology, Humboldt University of Berlin, Rudower Chaussee 18, 12489 Berlin, Germany, $^{98}$Epiq Solutions, Schaumburg, IL 60173, USA, $^{99}$Max Planck Institute for Software Systems, Saarbr\"ucken, Germany, $^{100}$Department of Psychology and Brain Sciences, Dartmouth College, Hanover, NH 03755, USA, $^{101}$Jacobs School of Engineering, University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92161, USA