### 2019-07-03 scipy-1.0/covlett.tex
Update scipy-1.0/covlett.tex
Co-Authored-By: Alessandro Pietro Bardelli <apbard@users.noreply.github.com>

% based on template:
% https://tex.stackexchange.com/questions/15532/looking-for-cover-letter-template

\documentclass[10pt,stdletter,dateno]{newlfm}
\usepackage{kpfonts}
\usepackage{url}

\widowpenalty=1000
\clubpenalty=1000

\newlfmP{headermarginskip=10pt}
\newlfmP{sigsize=50pt}
\newlfmP{dateskipafter=20pt}
%\newlfmP{addrfromphone}
%\newlfmP{addrfromemail}
%\PhrPhone{Phone}
%\PhrEmail{Email}

\namefrom{Ralf Gommers,
          Matt Haberland,
          and Tyler Reddy}
\addrfrom{%
    \today\\[10pt]
    Quansight Labs,\\
    The Netherlands\\
    \\
    BioResource and Agricultural Engineering,\\
    California Polytechnic State University,\\
    San Luis Obispo, CA, 93407, USA\\
    \\
    CCS-7 Applied Computer Science Group,\\
    Los Alamos National Laboratory,\\
    Los Alamos, NM, 87545, USA
}

\addrto{%
Prof. Syma Khalid,\\
School of Chemistry,\\
University of Southampton,\\
Highfield Campus,\\
Southampton,\\
SO17 1BJ, U.K.}

\greetto{Dear Prof. Khalid,}
\closeline{Sincerely,}
\begin{document}
\begin{newlfm}

We are submitting the enclosed manuscript describing a milestone
1.0 release of SciPy for consideration to be published in
\emph{Scientific Reports}. SciPy is the fundamental scientific 
computing library in the Python programming language. 
Hundreds of experts have contributed to our library, which
has been downloaded millions of times and has over 100,000 dependent
code repositories.

We have specifically chosen \emph{Scientific Reports} because
the breadth of acceptable topics matches the variety of algorithms
our library provides. High-profile manuscripts citing SciPy range
from the biosciences to recent reports on gravitational waves and
black holes. These citations target a website, and we now think
it is well past time that our mature library has a peer-reviewed
journal article describing it.

We suggest the following three peer reviewers as suitable candidates,
and we do not have any specific reviewer exclusion requests:

Thank you for your consideration.  We look forward to
hearing from you.

\end{newlfm}
\end{document}