### 2019-07-06 scipy-1.0/core-dev.tex
Updated author list

\author[1]{Pauli Virtanen}
\author[2]{Ralf Gommers}
\author[3,4,5,6,2]{Travis E. Oliphant}
\author[7,8]{Matt Haberland}
\author[9]{Tyler Reddy}
\author[10]{David Cournapeau}
\author[11]{Evgeni Burovski}
\author[12,13]{Pearu Peterson}
\author[10]{Warren Weckesser}
\author[14]{Jonathan Bright}
\author[15]{St\'efan J. van der Walt}
\author[16]{Matthew Brett}
\author[10]{Joshua Wilson}
\author[15,17]{K. Jarrod Millman}
\author[18]{Nikolay Mayorov}
\author[19]{Andrew R. J. Nelson}
\author[5]{Eric Jones}
\author[5]{Robert Kern}
\author[20]{Eric Larson}
\author[21]{CJ Carey}
\author[10]{\.{I}lhan Polat}
\author[22]{Yu Feng}
\author[23]{Eric W. Moore}
\author[24]{Jake VanderPlas}
\author[10]{Denis Laxalde}
\author[10]{Josef Perktold}
\author[25]{Robert Cimrman}
\author[26]{Ian Henriksen}
\author[10]{E. A. Quintero}
\author[10]{Charles R Harris}
\author[27]{Anne M. Archibald}
\author[28]{Ant\^{o}nio H. Ribeiro}
\author[29]{Fabian Pedregosa}
\author[ ]{SciPy 1.0 Contributors}
\affil[1]{University of Jyv\"askyl\"a, FI-40014 University of Jyv\"askyl\"a, Finland}
\affil[2]{Quansight LLC, Austin, TX, USA}
\affil[3]{Ultrasound Imaging, Mayo Clinic, Rochester, MN 55902, USA}
\affil[4]{Electrical Engineering, Brigham Young University, Provo, UT 84604, USA}
\affil[5]{Enthought, Inc., Austin, TX 78701, USA}
\affil[6]{Anaconda Inc, Austin, TX 78701, USA}
\affil[7]{BioResource and Agricultural Engineering Department, California Polytechnic State University, San Luis Obispo, CA 93407, USA}
\affil[8]{Department of Mathematics, University of California, Los Angeles, CA 90025, USA}
\affil[9]{Los Alamos National Laboratory, Los Alamos, NM 87544}
\affil[10]{Independent Researcher}
\affil[11]{Higher School of Economics, Myasnitskaya 20, Moscow 101000 Russia}
\affil[12]{Independent Researcher, Saue, Estonia}
\affil[13]{Department of Mechanics and Applied Mathematics, Institute of Cybernetics at Tallinn Technical University, Akadeemia Rd 21 Tallinn, Estonia}
\affil[14]{Independent Researcher, NY, USA}
\affil[15]{Berkeley Institute for Data Science, University of California, Berkeley, CA 94720, USA}
\affil[16]{School of Psychology, University of Birmingham, Edgbaston, Birmigham B15 2TT, UK}
\affil[17]{Division of Biostatistics, University of California, Berkeley, CA 94720, USA}
\affil[18]{WayRay LLC, Skolkovo Innovation Center, Moscow, 143026, Russia}
\affil[19]{Australian Nuclear Science and Technology Organisation, Locked Bag 2001, Kirrawee DC, NSW 2232, Australia}
\affil[20]{Institute for Learning and Brain Sciences, University of Washington, Seattle, WA 98195, USA}
\affil[21]{College of Information and Computing Sciences, University of Massachusetts Amherst, Amherst, MA 01002, USA}
\affil[22]{Berkeley Center for Cosmological Physics, University of California, Berkeley, CA 94720, USA}
\affil[23]{Bruker Biospin Corp., 15 Fortune Drive, Billerica, MA 01821}
\affil[24]{University of Washington, Seattle, WA 98195, USA}
\affil[25]{New Technologies - Research Centre, University of West Bohemia, Univerzitní 8, 306 14, Plzeň, Czech Republic}
\affil[26]{Oden Institute for Computational Engineering and Sciences, The University of Texas at Austin, Austin, Texas 78712, USA}
\affil[27]{Anton Pannekoek Institute, P.O. Box 94249, 1090 GE Amsterdam, The Netherlands}
\affil[28]{Graduate Program in Electrical Engineering, Universidade Federal de Minas Gerais, 6627 Av. Antonio Carlos 31270-901 Belo Horizonte, Brazil}
\affil[29]{Google, Inc., 355 Main St., Cambridge, MA 02142, USA}
\affil[*]{scipy.articles@gmail.com}