### 2019-07-06 scipy-1.0/consortium.tex

{\bf
Aditya Vijaykumar$^{30,31}$, Alessandro Pietro Bardelli$^{10}$, Alex Rothberg$^{10}$, Andreas Hilboll$^{32}$, Andreas Kloeckner$^{33}$, Anthony Scopatz$^{2}$, Antony Lee$^{34}$, Ariel Rokem$^{35}$, C. Nathan Woods$^{9}$, Chad Fulton$^{36}$, Charles Masson$^{37}$, Christian Brueffer$^{38}$, Christian H\"aggstr\"om$^{39}$, Clark Fitzgerald$^{40}$, David A. Nicholson$^{41}$, David R Hagen$^{42}$, Dmitrii V. Pasechnik$^{43}$, Emanuele Olivetti$^{44}$, Eric Martin$^{10}$, Eric Wieser$^{45}$, Fabrice Silva$^{46}$, Felix Lenders$^{47,48,49}$, Florian Wilhelm$^{50}$, G. Young$^{10}$, Gavin A. Price$^{51}$, Gert-Ludwig Ingold$^{52}$, Gregory E. Allen$^{53}$, Gregory R. Lee$^{54,55}$, Herv\'e Audren$^{56}$, Irvin Probst$^{57}$, J\"org P. Dietrich$^{58,59}$, Jacob Silterra$^{60}$, James T Webber$^{61}$, Janko Slavi\v{c}$^{62}$, Joel Nothman$^{63}$, Johannes Buchner$^{64,65}$, Johannes Kulick$^{66}$, Johannes L. Sch\"{o}nberger$^{10}$, Jos\'e Vin\'icius de Miranda Cardoso$^{67}$, Joscha Reimer$^{68}$, Joseph Harrington$^{69}$, Juan Luis Cano Rodr'iguez$^{70}$, Juan Nunez-Iglesias$^{71}$, Justin Kuczynski$^{72}$, Kevin Tritz$^{73}$, Martin Thoma$^{74}$, Matthias K\"ummerer$^{75}$, Maximilian Bolingbroke$^{76}$, Michael Tartre$^{77}$, Mikhail Pak$^{78}$, Nathaniel J. Smith$^{10}$, Nikolai Nowaczyk$^{10}$, Nikolay Shebanov$^{79}$, Oleksandr Pavlyk$^{80}$, Paul van Mulbregt$^{29}$, Per A. Brodtkorb$^{81}$, Perry Lee$^{10}$, Robert T. McGibbon$^{82}$, Roman Feldbauer$^{83}$, Sam Lewis$^{84}$, Sam Tygier$^{85}$, Scott Sievert$^{86}$, Sebastiano Vigna$^{87}$, Sourav Singh$^{88}$, Stefan Peterson$^{10}$, Surhud More$^{89,90}$, Tadeusz Pudlik$^{91}$, Takuya Oshima$^{92}$, Thomas J. Pingel$^{93}$, Thomas P. Robitaille$^{94}$, Thomas Spura$^{10}$, Thouis R. Jones$^{95}$, Tim Cera$^{10}$, Tim Leslie$^{10}$, Tiziano Zito$^{96}$, Tom Krauss$^{97}$, Utkarsh Upadhyay$^{98}$, Yaroslav O. Halchenko$^{99}$, Yoshiki V\'azquez-Baeza$^{100}$
}
\newline
\hfill \break
$^{30}$International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bengaluru 560089, India, $^{31}$Department of Physics, Birla Institute of Technology and Science, Pilani 330331, India, $^{32}$Institute of Environmental Physics, University of Bremen, Bremen, Germany, $^{33}$Department of Computer Science, University of Illinois at Urbana-Champaign, 201 N. Goodwin Ave, Urbana, IL 616801, USA, $^{34}$Laboratoire Photonique, Num\'erique et Nanosciences UMR 5298, Universit\'e de Bordeaux, Institut d'Optique Graduate School, CNRS, 33400 Talence, France, $^{35}$The University of Washington eScience Institute, The University of Washington, Seattle, WA 98105, USA, $^{36}$Federal Reserve Board of Governors,  20th and Constitution Ave. NW, Washington, DC 20551, U.S.A., $^{37}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, $^{38}$Department of Clinical Sciences Lund, Lund University, Lund, Sweden, $^{39}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, $^{40}$Statistics Department, University of California - Davis, Davis, CA 95616, USA, $^{41}$Emory University, Atlanta, GA 30322, USA, $^{42}$Applied BioMath, 561 Virginia Rd Ste 220, Concord, MA 01742, USA, $^{43}$Department of Computer Science, University of Oxford, Wolfson Building, Parks Road, OX1 3QD Oxford, UK, $^{44}$NeuroInformatics Laboratory, Bruno Kessler Foundation, Trento 38123, Italy, $^{45}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, $^{46}$Aix Marseille Universit\'e., CNRS, Centrale Marseille, LMA, Marseille, France, $^{47}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, $^{48}$ABB Corporate Research, ABB AG, Wallstadter Str. 59 68526 Ladenburg, Germany, $^{49}$Institut f\"ur Mathematische Optimierung, Technische Universit\"at Carolo-Wilhelmina zu Braunschweig, Universit\"atsplatz 2 38106 Braunschweig, Germany, $^{50}$Independent Researcher, Cologne, Germany, $^{51}$Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA, $^{52}$Institut f{\"u}r Physik, Universit{\"a}t Augsburg, Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany, $^{53}$Applied Research Laboratories, The University of Texas at Austin, P.O. Box 8029 Austin, TX 78713, USA, $^{54}$Department of Radiology, School of Medicine, University of Cincinnati, Cincinnati, OH 45221, USA, $^{55}$Department of Radiology, Cincinnati Children's Hospital Medical Center, Cincinnati, OH 45229, USA, $^{56}$Ascent Robotics Inc., 1-6-10 Hiroo Shibuya Tokyo, Japan, $^{57}$ENSTA Bretagne, 2 rue François Verny, Brest, France, $^{58}$Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 1, 81679 M\"unchen, Germany, $^{59}$Excellence Cluster Universe, Boltzmannstr. 2, 85748 Garching b. M\"unchen, Germany, $^{60}$Independent Researcher, USA, $^{61}$Data Sciences, Chan Zuckerberg Biohub, San Francisco, CA 94158, USA, $^{62}$Faculty of Mechanical Engineering, University of Ljubljana, A\v{s}ker\v{c}eva 6 1000 Ljubljana, Slovenia-EU, $^{63}$Sydney Informatics Hub, The University of Sydney, NSW 2006, Australia, $^{64}$Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'{o}lica de Chile, Casilla 306, Santiago 22, Chile, $^{65}$Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85741 Garching, Germany, $^{66}$University of Stuttgart, Machine Learning and Robotics Lab, Universitätsstr. 38, 70569 Stuttgart, Germany, $^{67}$Department of Electrical Engineering, Universidade Federal de Campina Grande, 58429900 PB, Brazil, $^{68}$Department of Computer Science, Kiel University, 24098 Kiel, Germany, $^{69}$Planetary Sciences Group and Florida Space Institute and Department of Physics, University of Central Florida, 4111 Libra Drive Orlando Florida 32816-2385, USA, $^{70}$Independent Researcher, Spain, $^{71}$Monash Micro Imaging, Monash University, Clayton VIC 3800, Australia, $^{72}$Department of Molecular, Cellular, and Developmental Biology, University of Colorado, Boulder, Boulder, CO 80303, USA, $^{73}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N. Charles Street, Baltimore, MD 21218, USA, $^{74}$Independent Researcher, Munich, Germany, $^{75}$University of T\"ubingen, T\"ubingen, Germany, $^{76}$Independent Researcher, United Kingdom, $^{77}$Two Sigma Investments, 100 Avenue of the Americas 16th Floor, New York, NY 10013, USA, $^{78}$Department of Mechanical Engineering, Technical University of Munich, Boltzmannstr. 15, 85748 Garching, Germany, $^{79}$Independent Researcher, Berlin, Germany, $^{80}$Intel Corp., Austin, TX 78746, USA, $^{81}$Independent Researcher, Horten, Norway, $^{82}$D. E. Shaw Research, New York, NY 10036, USA, $^{83}$Division of Computational Systems Biology, University of Vienna, Althanstra{\ss}e 14, 1090 Vienna, Austria, $^{84}$Independent Researcher, Melbourne, Australia, $^{85}$School of Physics and Astronomy, University of Manchester, Oxford Road, M13 9PL, UK, $^{86}$Electrical and Computer Engineering, University of Wisconsin--Madison, 1415 Engineering Drive Madison, WI, USA, $^{87}$Dipartimento di Informatica, Universit\`a degli Studi di Milano, via Celoria 18, 20133 Milano MI, Italy, $^{88}$Department of Computer Engineering, Vishwakarma Institute of Information Technology, Pune, India, $^{89}$Inter-University Centre for Astronomy and Astrophysics, Ganeshkhind, Pune, 411007, India, $^{90}$Kavli Institute for the Physics and Mathematics of the Universe, Kashiwanoha 5-1-5, Kashiwa-shi, 2778583, Japan, $^{91}$Waymo LLC, Mountain View, CA 94043, USA, $^{92}$Faculty of Engineering, Niigata University, 8050 Ikarashi-Ninocho, Nishi-ku, Niigata, 9502181, Japan, $^{93}$Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA, $^{94}$Aperio Software, Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom, $^{95}$Broad Institute, Cambridge, MA 02142, USA, $^{96}$Department of Psychology, Humboldt University of Berlin, Rudower Chaussee 18, 12489 Berlin, Germany, $^{97}$Epiq Solutions, Schaumburg, IL 60173, USA, $^{98}$Max Planck Institute for Software Systems, Saarbr\"ucken, Germany, $^{99}$Department of Psychology and Brain Sciences, Dartmouth College, Hanover, NH 03755, USA, $^{100}$Jacobs School of Engineering, University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92161, USA