### 2019-09-19 scipy-1.0/consortium.tex

{\bfseries
Aditya Vijaykumar$^{32,33}$, 
Alessandro Pietro Bardelli$^{10}$, 
Alex Rothberg$^{10}$, 
Andreas Hilboll$^{34}$, 
Andreas Kloeckner$^{35}$, 
Anthony Scopatz$^{2}$, 
Antony Lee$^{36}$, 
Ariel Rokem$^{37}$, 
C. Nathan Woods$^{9}$, 
Chad Fulton$^{38}$, 
Charles Masson$^{39}$, 
Christian H\"aggstr\"om$^{40}$, 
Clark Fitzgerald$^{41}$, 
David A. Nicholson$^{42}$, 
David R Hagen$^{43}$, 
Dmitrii V. Pasechnik$^{44}$, 
Emanuele Olivetti$^{45}$, 
Eric Martin$^{10}$, 
Eric Wieser$^{46}$, 
Fabrice Silva$^{47}$, 
Felix Lenders$^{48,49,50}$, 
Florian Wilhelm$^{51}$, 
G. Young$^{10}$, 
Gavin A. Price$^{52}$, 
Gert-Ludwig Ingold$^{53}$, 
Gregory E. Allen$^{54}$, 
Gregory R. Lee$^{55,56}$, 
Herv\'e Audren$^{57}$, 
Irvin Probst$^{58}$, 
J\"org P. Dietrich$^{59,60}$, 
Jacob Silterra$^{61}$, 
James T Webber$^{62}$, 
Janko Slavi\v{c}$^{63}$, 
Joel Nothman$^{64}$, 
Johannes Buchner$^{65,66}$, 
Johannes Kulick$^{67}$, 
Johannes L. Sch\"{o}nberger$^{10}$, 
Jos\'e Vin\'icius de Miranda Cardoso$^{68}$, 
Joscha Reimer$^{69}$, 
Joseph Harrington$^{70}$, 
Juan Luis Cano Rodr\'iguez$^{71}$, 
Juan Nunez-Iglesias$^{72}$, 
Justin Kuczynski$^{73}$, 
Kevin Tritz$^{74}$, 
Martin Thoma$^{75}$, 
Matthew Newville$^{76}$, 
Matthias K\"ummerer$^{77}$, 
Maximilian Bolingbroke$^{78}$, 
Michael Tartre$^{79}$, 
Mikhail Pak$^{80}$, 
Nathaniel J. Smith$^{10}$, 
Nikolai Nowaczyk$^{10}$, 
Nikolay Shebanov$^{81}$, 
Oleksandr Pavlyk$^{82}$, 
Per A. Brodtkorb$^{83}$, 
Perry Lee$^{10}$, 
Robert T. McGibbon$^{84}$, 
Roman Feldbauer$^{85}$, 
Sam Lewis$^{86}$, 
Sam Tygier$^{87}$, 
Scott Sievert$^{88}$, 
Sebastiano Vigna$^{89}$, 
Stefan Peterson$^{10}$, 
Surhud More$^{90,91}$, 
Tadeusz Pudlik$^{92}$, 
Takuya Oshima$^{93}$, 
Thomas J. Pingel$^{94}$, 
Thomas P. Robitaille$^{95}$, 
Thomas Spura$^{10}$, 
Thouis R. Jones$^{96}$, 
Tim Cera$^{10}$, 
Tim Leslie$^{10}$, 
Tiziano Zito$^{97}$, 
Tom Krauss$^{98}$, 
Utkarsh Upadhyay$^{99}$, 
Yaroslav O. Halchenko$^{100}$, 
Yoshiki V\'azquez-Baeza$^{101}$
}
\newline
\hfill \break
$^{32}$International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bengaluru 560089, India, 
$^{33}$Department of Physics, Birla Institute of Technology and Science, Pilani 330331, India, 
$^{34}$Institute of Environmental Physics, University of Bremen, Bremen, Germany, 
$^{35}$Department of Computer Science, University of Illinois at Urbana-Champaign, 201 N. Goodwin Ave, Urbana, IL 616801, USA, 
$^{36}$Laboratoire Photonique, Num\'erique et Nanosciences UMR 5298, Universit\'e de Bordeaux, Institut d'Optique Graduate School, CNRS, 33400 Talence, France, 
$^{37}$The University of Washington eScience Institute, The University of Washington, Seattle, WA 98105, USA, 
$^{38}$Federal Reserve Board of Governors,  20th and Constitution Ave. NW, Washington, DC 20551, U.S.A., 
$^{39}$Datadog Inc., 620 8th Avenue New York, NY 10018, USA, 
$^{40}$HQ, Orexplore, Torshamnsgatan 30B, Stockholm, Sweden, 
$^{41}$Statistics Department, University of California - Davis, Davis, CA 95616, USA, 
$^{42}$Emory University, Atlanta, GA 30322, USA, 
$^{43}$Applied BioMath, 561 Virginia Rd Ste 220, Concord, MA 01742, USA, 
$^{44}$Department of Computer Science, University of Oxford, Wolfson Building, Parks Road, OX1 3QD Oxford, UK, 
$^{45}$NeuroInformatics Laboratory, Bruno Kessler Foundation, Trento 38123, Italy, 
$^{46}$Department of Engineering, University of Cambridge, Trumpington St, Cambridge CB2 1PZ, UK, 
$^{47}$Aix Marseille Univ, CNRS, Centrale Marseille, LMA, Marseille, France, 
$^{48}$Interdisciplinary Center for Scientific Computing (IWR), Heidelberg University, Im Neuenheimer Feld 205 69120 Heidelberg, Germany, 
$^{49}$ABB Corporate Research, ABB AG, Wallstadter Str. 59 68526 Ladenburg, Germany, 
$^{50}$Institut f\"ur Mathematische Optimierung, Technische Universit\"at Carolo-Wilhelmina zu Braunschweig, Universit\"atsplatz 2 38106 Braunschweig, Germany, 
$^{51}$Independent Researcher, Cologne, Germany, 
$^{52}$Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA, 
$^{53}$Institut f{\"u}r Physik, Universit{\"a}t Augsburg, Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany, 
$^{54}$Applied Research Laboratories, The University of Texas at Austin, P.O. Box 8029 Austin, TX 78713, USA, 
$^{55}$Department of Radiology, School of Medicine, University of Cincinnati, Cincinnati, OH 45221, USA, 
$^{56}$Department of Radiology, Cincinnati Children's Hospital Medical Center, Cincinnati, OH 45229, USA, 
$^{57}$Ascent Robotics Inc., 1-6-10 Hiroo Shibuya Tokyo, Japan, 
$^{58}$ENSTA Bretagne, 2 rue François Verny, Brest, France, 
$^{59}$Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 1, 81679 M\"unchen, Germany, 
$^{60}$Excellence Cluster Universe, Boltzmannstr. 2, 85748 Garching b. M\"unchen, Germany, 
$^{61}$Independent Researcher, USA, 
$^{62}$Data Sciences, Chan Zuckerberg Biohub, San Francisco, CA 94158, USA, 
$^{63}$Faculty of Mechanical Engineering, University of Ljubljana, A\v{s}ker\v{c}eva 6 1000 Ljubljana, Slovenia-EU, 
$^{64}$Sydney Informatics Hub, The University of Sydney, NSW 2006, Australia, 
$^{65}$Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'{o}lica de Chile, Casilla 306, Santiago 22, Chile, 
$^{66}$Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85741 Garching, Germany, 
$^{67}$University of Stuttgart, Machine Learning and Robotics Lab, Universitätsstr. 38, 70569 Stuttgart, Germany, 
$^{68}$Department of Electrical Engineering, Universidade Federal de Campina Grande, 58429900 PB, Brazil, 
$^{69}$Department of Computer Science, Kiel University, 24098 Kiel, Germany, 
$^{70}$Planetary Sciences Group and Florida Space Institute and Department of Physics, University of Central Florida, 4111 Libra Drive Orlando Florida 32816-2385, USA, 
$^{71}$Independent Researcher, Spain, 
$^{72}$Monash Micro Imaging, Monash University, Clayton VIC 3800, Australia, 
$^{73}$Department of Molecular, Cellular, and Developmental Biology, University of Colorado, Boulder, Boulder, CO 80303, USA, 
$^{74}$Department of Physics and Astronomy, Johns Hopkins University, 3400 N. Charles Street, Baltimore, MD 21218, USA, 
$^{75}$Independent Researcher, Munich, Germany, 
$^{76}$Center for Advanced Radiation Sources, The University of Chicago, Chicago, IL 60637, USA, 
$^{77}$University of T\"ubingen, T\"ubingen, Germany, 
$^{78}$Independent Researcher, United Kingdom, 
$^{79}$Two Sigma Investments, 100 Avenue of the Americas 16th Floor, New York, NY 10013, USA, 
$^{80}$Department of Mechanical Engineering, Technical University of Munich, Boltzmannstr. 15, 85748 Garching, Germany, 
$^{81}$Independent Researcher, Berlin, Germany, 
$^{82}$Intel Corp., Austin, TX 78746, USA, 
$^{83}$Independent Researcher, Horten, Norway, 
$^{84}$D. E. Shaw Research, New York, NY 10036, USA, 
$^{85}$Division of Computational Systems Biology, Department of Microbiology and Ecosystem Science, University of Vienna, Althanstra{\ss}e 14, 1090 Vienna, Austria, 
$^{86}$Independent Researcher, Melbourne, Australia, 
$^{87}$School of Physics and Astronomy, University of Manchester, Oxford Road, M13 9PL, UK, 
$^{88}$Electrical and Computer Engineering, University of Wisconsin--Madison, 1415 Engineering Drive Madison, WI, USA, 
$^{89}$Dipartimento di Informatica, Universit\`a degli Studi di Milano, via Celoria 18, 20133 Milano MI, Italy, 
$^{90}$Inter-University Centre for Astronomy and Astrophysics, Ganeshkhind, Pune, 411007, India, 
$^{91}$Kavli Institute for the Physics and Mathematics of the Universe, Kashiwanoha 5-1-5, Kashiwa-shi, 2778583, Japan, 
$^{92}$Waymo LLC, Mountain View, CA 94043, USA, 
$^{93}$Faculty of Engineering, Niigata University, 8050 Ikarashi-Ninocho, Nishi-ku, Niigata, 9502181, Japan, 
$^{94}$Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA, 
$^{95}$Aperio Software, Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom, 
$^{96}$Broad Institute, Cambridge, MA 02142, USA, 
$^{97}$Department of Psychology, Humboldt University of Berlin, Rudower Chaussee 18, 12489 Berlin, Germany, 
$^{98}$Epiq Solutions, Schaumburg, IL 60173, USA, 
$^{99}$Max Planck Institute for Software Systems, Kaiserslautern, Germany, 
$^{100}$Department of Psychology and Brain Sciences, Dartmouth College, Hanover, NH 03755, USA, 
$^{101}$Jacobs School of Engineering, University of California San Diego, 9500 Gilman Drive, La Jolla, CA 92161, USA